\Title

### 2013-01-28 synmut.tex
text revisited around new figure

\begin{figure}
\begin{center}
\subfloat{\includegraphics[width=\linewidth]{simulations_graduallydel.pdf}
\label{fig:simfixpvar}}\\
\subfloat{\includegraphics[width=\linewidth]{simulations_syn.pdf}
\label{fig:simsfig}}
\caption{(Panel A) The depression in $P_\text{fix}$ depends on the deleterious
effect size of the synonymous alleles (in HIV data is about $-0.2$).
This parameter also reduces synonymous
diversity, measured by the number of mutations at frequencies $0.25 < \nu < 0.75$,
normalized by the number of possible mutations (first inset).
(Panel B) To assess the parameter space that affects synonymous fixation and
diversity, we run many simulations with random parameters for deleterious effect
size, fraction of deleterious synonymous sites, average size of escape
mutations (color, blue ($10^{-2.5}$) to red ($10^{-1.5}$)), and rate of
introduction of new epitopes (marker size, from $10^{-3}$ to $10^{-2}$ per
generation). Mostly simulations with small deleterious effects and relatively
high deleterious fractions of synonymous mutations reproduce synonymous HIV-like
fixation probability and diversity.}
\label{fig:simheat}
\end{center}
\end{figure}

@@

The first result of the simulations is that genetic draft can indeed bring weakly
deleterious mutations to high frequencies and result in a dependence of the
fixation probability on initial frequency that is compatible with observations
(see \FIG{simfixpvar}).
The fixation probability of synonymous alleles decreases from the neutral
expectation to zero as their fitness effect worsens (\FIG{simfixpvar},
from $-0.0003$ in blue to $-0.003$ in red). Contextually,
the number of alleles that reach high frequencies, which is a proxy for
synonymous diversity, is reduced as well (\FIG{simfixpvar}, top inset).
To assess the reduction in fixation probability quantitatively, we calculathe
the area between the measured fixation probability and the diagonal, which is
the neutral expectation (\FIG{simfixpvar}, lower inset). This area spans a range
between $-0.5$ (no fixation at all), zero (neutral-like curve), and $+0.5$ (always
fixed). The HIV data correspond to an area of approximately $-0.2$ for synonymous
changes, and essentially zero for nonsynonymous changes.

Among all simulations, we select the one that show a large depression in
fixation probability of synonymous mutations but, simulteneously, a moderately
high synonymous diversity, quantified by the probability $P_\text{interm}$ of a
derived allele to be at intermediate frequencies $0.25 < \nu < 0.75$. HIV data show
$A_\text{syn} \sim -0.2$ and $P_\text{interm} \sim 0.005$; simulations are gated
such that $A_\text{syn} < -0.15$ and $0.0025 < P_\text{interm} < 0.010$, hence
allowing an uncertainty on mutation rates or generation times of about twofold.
This selection imposes a strong
constraint on the parameters, in particular on the ones related to synonymous
changes, i.e. their deleterious fraction and effect size (see \FIG{simsfig}).
A high fraction ($\gtrsim 0.8$) of only slightly deleterious synonymous sites,
with effect size $|s_d| \sim 0.002$, is required. This result fits well
the expectation based on the fixation/extinction times above (see \FIG{fixp1}).
Moreover, it corresponds to {\it a priori} theoretical expectations:
(a) since high frequency bins are enriched in neutral mutations (it is easier for
them to hitchhike), the hallmark of deleterious mutations will be visible there
only if they are pervasive; (b) genetic draft stops working if the
deleterious effect size becomes comparable to the adaptive effect size of escape
mutations, because the double mutant has little or no fitness advantage; (c)
mutations with a deleterious effect weaker than approximately $0.001$ are
indistinguishable from neutral ones and are drafted seamlessly to any frequency.

>>>>

\begin{figure}
\begin{center}
\subfloat{\includegraphics[width=\linewidth]{simulations_graduallydel.pdf}
\label{fig:simfixpvar}}\\
\subfloat{\includegraphics[width=\linewidth]{simulations_syn.pdf}
\label{fig:simsfig}}
\caption{(Panel A) The depression in $P_\text{fix}$ depends on the deleterious
effect size of the synonymous alleles. This parameter also reduces synonymous
diversity, measured by probability of a derived allele to be found at
intermediate frequencies $P_\text{interm}$ (first inset).
(Panel B) To assess the parameter space that affects synonymous fixation and
diversity, we run many simulations with random parameters for deleterious effect
size, fraction of deleterious synonymous sites, average size of escape
mutations (color, blue ($10^{-2.5}$) to red ($10^{-1.5}$)), and rate of
introduction of new epitopes (marker size, from $10^{-3}$ to $10^{-2}$ per
generation). Mostly simulations with small deleterious effects and relatively
high deleterious fractions of synonymous mutations reproduce synonymous HIV-like
fixation probability and diversity.}
\label{fig:simheat}
\end{center}
\end{figure}

@@

The first result of the simulations is that genetic draft can indeed bring weakly
deleterious mutations to high frequencies and result in a dependence of the
fixation probability on initial frequency that is compatible with observations
(see \FIG{simfixpvar}).
To assess the reduction in fixation probability quantitatively, we calculate
the area between the measured fixation probability and the diagonal, which is
the neutral expectation (\FIG{simfixpvar}, lower inset). This area spans a range
between $-0.5$ (no fixation at all), zero (neutral-like curve), and $+0.5$ (always
fixed). Furthermore, we quantify synonymous diversity via the probability
$P_\text{interm}$ of a synonymous derived allele to be at intermediate frequencies
$0.25 < \nu < 0.75$ (\FIG{simfixpvar}, top inset).
The HIV data correspond to an area $A_\text{syn} \sim -0.2$  for synonymous
changes and $A_\text{nonsyn} \sim 0$ for nonsynonymous changes,
and a synonymous diversity $P_\text{interm} \sim 0.005$.
In the three simulations shown, the fixation probability of synonymous alleles
decreases from the neutral expectation ($A_\text{syn} \sim 0$) to zero
($A_\text{syn} \sim -0.5$) as their fitness effect
worsens; the synonymous diversity plummets as well, as it is harder for
deleterious mutations to reach the frequency threshold of 0.25
(\FIG{simfixpvar}, from $-0.0003$ in blue to $-0.003$ in red).

Among all simulations, we select the ones that show HIV-like $A_\text{syn}$ and
$P_\text{interm}$, i.e. a large depression in
fixation probability of synonymous mutations but, simulteneously, a moderately
high synonymous diversity: simulations are gated such that $A_\text{syn} < -0.15$
and $0.0025 < P_\text{interm} < 0.010$, hence allowing an uncertainty on mutation
rates or generation times of about twofold. This selection imposes a strong
constraint on the parameters, in particular on the ones related to synonymous
changes, i.e. their deleterious fraction and effect size (see \FIG{simsfig}).
A high fraction ($\gtrsim 0.8$) of only slightly deleterious synonymous sites,
with effect size $|s_d| \sim 0.002$, is required. This result fits well
the expectation based on the fixation/extinction times above (see \FIG{fixp1}).
Moreover, it corresponds to {\it a priori} theoretical expectations:
(a) since high frequency bins are enriched in neutral mutations (it is easier for
them to hitchhike), the hallmark of deleterious mutations will be visible there
only if they are pervasive; (b) genetic draft stops working if the
deleterious effect size becomes comparable to the adaptive effect size of escape
mutations, because the double mutant has little or no fitness advantage; (c)
mutations with a deleterious effect weaker than approximately $0.001$ are
indistinguishable from neutral ones and are drafted seamlessly to any frequency.
